---
title: 'Interactive Charts: htmlwidgets'
author: "Vivek H. Patil"
date: "November 6, 2015"
output: html_document
---

<style>
.title-slide hgroup h1 {color: red;}
h2 {color: red;}
slides > slide:not(.nobackground):after {
  content: '';
}
</style>

```{r global_options, include=FALSE}
library(knitr)
library(printr)
opts_chunk$set(warning=FALSE, message=FALSE)
```

## This presentation

* Presentation: [http://patilv.com/INRUG-InteractiveCharts](http://patilv.com/INRUG-InteractiveCharts)
* Single page version of this presentation: [http://patilv.com/INRUG-InteractiveCharts/allinone.html](http://patilv.com/INRUG-InteractiveCharts/allinone.html)
* Created using R markdown and R-Studio
* GitHub repository for source : [https://github.com/patilv/INRUG-InteractiveCharts](https://github.com/patilv/INRUG-InteractiveCharts)

## htmlwidgets, rCharts

* `htmlwidgets`: Ramnath Vaidyanathan, Kenton Russell, and RStudio
* `rCharts` (a transition to 2.0 using `htmlwidgets` is underway): Ramnath Vaidyanathan

* Provide the ability, from within R, to create, customize and publish interactive JavaScript (JS) visualizations
* JS is a computer programming language that can create interactivity using web browsers
* There are many pre-written JS libraries that make the process of creating visualizations easier. 

```{r eval=FALSE}

install.packages("htmlwidgets")
```

## Challenge in learning

* Different JS libraries developed independently - consistency in calls can at times be a challenge
* Customization for specific needs may require writing JS code

### Learning approach

* Follow examples, questions, and issues presented in different venues
* Stick to few JS libraries which provide many choices of charts and become very familiar with them. Don't forget that few have restrictions for commercial use.


## Few Resources

* `htmlwidgets`
    * [htmlwidgets' site](http://www.htmlwidgets.org/index.html)
    * [Timelyportfolio's challenge - A widget a week](http://www.buildingwidgets.com/)
    * [Ryan Hafen's gallery of htmlwidgets developed by different people](http://hafen.github.io/htmlwidgetsgallery/)
    * [Search GitHub](http://www.github.com) for `htmlwidgets`
* `rCharts` (A reboot to an `htmlwidgets` version is on the way)
    * [rCharts' website](http://rcharts.io/) and its [gallery of submissions from different people](http://rcharts.io/gallery/).
    * [rCharts on GitHub and the issues posted there](https://github.com/ramnathv/rCharts) and on 

## A sampling of the several htmlwidget packages 
|   |   |
|---|---|
| Interactive tables | [DT](https://rstudio.github.io/DT/), [formattable](http://renkun.me/formattable/),[datacomb](https://github.com/cmpolis/datacomb), [rpivotTable](https://github.com/smartinsightsfromdata/rpivotTable),  [rhandsontable](http://jrowen.github.io/rhandsontable/), [D3TableFilter](https://github.com/ThomasSiegmund/D3TableFilter),  [sortableR](https://github.com/timelyportfolio/sortableR) |
| Heatmaps  | [d3heatmap](https://github.com/rstudio/d3heatmap)|
|Dimple|[rcdimple](https://github.com/timelyportfolio/rcdimple)|
| Tau Charts|[taucharts](https://github.com/hrbrmstr/taucharts)|
|Metrics Graphics|[metricsgraphics](http://hrbrmstr.github.io/metricsgraphics/)|
| Bokeh | [rbokeh](http://hafen.github.io/rbokeh/)|

## Few more sample packages 
|   |   |
|---|---|
|streamgraph|[streamgraph](http://hrbrmstr.github.io/streamgraph/)|
|Scatterplot matrices|[pairsD3](https://github.com/garthtarr/pairsD3), [scatterMatrixD3](https://github.com/jcizel/scatterMatrixD3)|
|Pan and Zoom graphs|[svgPanZoom](https://github.com/timelyportfolio/svgPanZoom)|
|Parallel Coordinate Charts|[parcoords](https://github.com/timelyportfolio/parcoords)|
|Sequences Sunburst|[sunburstR](http://www.buildingwidgets.com/blog/2015/7/2/week-26-sunburstr)|

## Example 1: Interactive table

```{r}
library(htmlwidgets)
suppressPackageStartupMessages(library(dplyr))
library(DT)# if (!require("DT")) install.packages('DT')
dt=datatable(iris, options=list(pageLength = 5))
saveWidget(dt,file = "dt.html",selfcontained = TRUE)
```
<iframe src="dt.html" width="800" height="600"></iframe>

## Example 2: Heatmap

```{r}
head(mtcars)
```

## Example 2 continued: Heatmap
```{r}
library(d3heatmap)#if (!require("d3heatmap")) devtools::install_github("rstudio/d3heatmap")
myheatmap=d3heatmap(mtcars,scale="column")
saveWidget(myheatmap,file = "myheatmap.html",selfcontained = TRUE)
```


## Example 2 continued: Heatmap

<iframe src="myheatmap.html" width="800" height="600"></iframe>

## Example 3: `rcdimple`: Many graph types available
[See examples of different types here.](http://patilv.com/rcdimpleexamples/)
```{r}
suppressPackageStartupMessages(library(rcdimple)) # demo(dimple) for tons of examples#if (!require("rcdimple")) devtools::install_github("timelyportfolio/rcdimple")
myirisplot=iris%>% dimple(x = "Sepal.Length",
  y = "Sepal.Width",
  groups = "Species",
  type = "bubble",width=600,height=380)%>% yAxis( overrideMin = 2)%>%
  add_legend()
saveWidget(myirisplot,file = "myirisplot.html",selfcontained = TRUE)
```

## Example 3 continued
<iframe src="myirisplot.html" width="800" height="600"></iframe>

## Example 4 grouped bar

```{r}
gearam=mtcars%>%group_by(gear,am)%>%summarise(count=length(gear))
gearam$am<-ifelse(gearam$am==0,yes= "Auto",no= "Manual")
gearam
```

## Example 4 grouped bar: `rcdimple`: Many graph types available
```{r}
mydimplebarplot=gearam%>%dimple(
    x = c("gear","am"), y = "count",
    groups = "am", type = "bar",width=600,height=390)%>%
   add_legend()
saveWidget(mydimplebarplot,file = "mydimplebarplot.html",selfcontained = TRUE)
```

## Example 4 continued
<iframe src="mydimplebarplot.html" width="800" height="600"></iframe>


##  Example 5: `taucharts`: Many graph types
```{r}
library(taucharts)#if (!require("taucharts")) devtools::install_github("hrbrmstr/taucharts")
tauirisplot=iris %>% tauchart(height=500,width=800)%>% 
  tau_point(x="Sepal.Length",y="Sepal.Width",color="Species")%>%
  tau_tooltip(c("Sepal.Length", "Sepal.Width", "Species"))%>% 
  tau_guide_y(auto_scale=FALSE)%>% 
  tau_guide_x(auto_scale=FALSE)%>%tau_legend()
saveWidget(tauirisplot,file = "tauirisplot.html",selfcontained = TRUE)
```

## Example 5 continued
<iframe src="tauirisplot.html" width="800" height="700"></iframe>

## Example 6: `taucharts`: Many graph types

```{r}
gearam$gear=as.factor(gearam$gear)
gearbartauplot=tauchart(gearam)%>%tau_bar("gear","count","am")%>%
  tau_tooltip(c("gear", "am", "count"))%>% tau_legend()
saveWidget(gearbartauplot,file = "gearbartauplot.html",selfcontained = TRUE)
```

## Example 6 continued
<iframe src="gearbartauplot.html" width="800" height="800"></iframe>

## Example 7: `metricsgraphics`: Many graph types available

Scatter and line plots (time-series data), grids/facets of graphs, and connected graphs
```{r}
#if (!require("metricsgraphics")) devtools::install_github("hrbrmstr/metricsgraphics")
library(metricsgraphics)
irismetricsplot=iris %>% mjs_plot(x=Sepal.Length,y=Sepal.Width) %>%
  mjs_point(color_accessor=Species, color_type="category")%>%
    mjs_labs(x="Sepal.Length", y="Sepal.Width")%>% 
    mjs_axis_y(min_y=2)
saveWidget(irismetricsplot,file = "irismetricsplot.html",selfcontained = TRUE)
```
## Example 7 continued
<iframe src="irismetricsplot.html" width="950" height="800"></iframe>

## Example 8: `rbokeh`: Many graph types available
[http://hafen.github.io/rbokeh/](http://hafen.github.io/rbokeh/): Bar missing
```{r}
library(rbokeh)#if (!require("rbokeh")) devtools::install_github("bokeh/rbokeh")
bokehirisplot=figure() %>%
  ly_points(Sepal.Length, Sepal.Width, data = iris,
    color = Species,
    hover = list(Sepal.Length, Sepal.Width))
saveWidget(bokehirisplot,file = "bokehirisplot.html",selfcontained = TRUE)
```


## Example 8 continued
<iframe src="bokehirisplot.html" width="800" height="850"></iframe>



## Example 9: Scatter plot matrix `pairsD3`

```{r}
#if (!require("pairsD3")) devtools::install_github("garthtarr/pairsD3")
library(pairsD3)
pairsd3plot=pairsD3(iris[,1:4],group=iris[,5])#%>%savePairs(file = 'iris.html')
saveWidget(pairsd3plot,file = "pairsd3plot.html",selfcontained = TRUE)
```

## Example 9 continued
<iframe src="pairsd3plot.html" width="800" height="600"></iframe>

## Example 10: Scatter plot matrix `scatterMatrixD3`

```{r}
#if (!require("scatterMatrixD3")) devtools::install_github("jcizel/scatterMatrixD3")
library(scatterMatrixD3)
scattermatplot=scatterMatrix(
   data = iris
)
saveWidget(scattermatplot,file = "scattermatplot.html",selfcontained = TRUE)
```

## Example 10 continued
<iframe src="scattermatplot.html" width="800" height="800"></iframe>

## Example 11 `streamgraph`
```{r eval=FALSE}
#if (!require("streamgraph")) devtools::install_github("hrbrmstr/streamgraph")
library(streamgraph)
library(babynames)
head(babynames,3)
maryplot=babynames %>%
  filter(grepl("^Mar", name)) %>%
  group_by(year, name) %>%
  tally(wt=n) %>%
  streamgraph("name", "n", "year")
saveWidget(maryplot,file = "maryplot.html",selfcontained = TRUE)
```

## Example 11 continued
<iframe src="maryplot.html" width="850" height="800"></iframe>

## Example 12 `svgPanZoom`

```{r}
#if (!require("svgPanZoom")) devtools::install_github("timelyportfolio/svgPanZoom")
library(svgPanZoom)
library(ggplot2)
mynewggplot=svgPanZoom(
  ggplot(iris,aes(Sepal.Length,Sepal.Width,color=Species))+geom_point()
)
saveWidget(mynewggplot,file = "mynewggplot.html",selfcontained = TRUE)
```



## Example 12 continued

<iframe src="mynewggplot.html" width="800" height="700"></iframe>

## Example 13 `parcoords`

```{r}
#if (!require("parcoords")) devtools::install_github("timelyportfolio/parcoords")
library(parcoords)
mymtcarsplot=parcoords(mtcars,brushMode = "1D-axes-multi")
saveWidget(mymtcarsplot,file = "mymtcarsplot.html",selfcontained = TRUE)
```


## Example 13 continued

<iframe src="mymtcarsplot.html" width="800" height="700"></iframe>


## Example 14 `sunburstr`
```{r}
#if (!require("sunburstR")) devtools::install_github("timelyportfolio/sunburstR")
library(sunburstR)
school=read.csv("school.csv")
school
```
## Example 14 continued

```{r}
school$combo=paste(school$School,school$Department,school$Specialization,sep="-")
school$combo=gsub("-$","",school$combo)
head(school,1)
myburst=sunburst(school[,c(5,4)])
saveWidget(myburst,file = "myburst.html",selfcontained = TRUE)
```



## Example 14 continued

<iframe src="myburst.html" width="800" height="800"></iframe>

## More packages

* `leaflet`: Mapping
* Network data and Graph Diagrams
    * mermaid and grViz: [DiagrammeR](http://rich-iannone.github.io/DiagrammeR/index.html) 
    * Vivagraph: [vivagRaph](https://github.com/keeganhines/vivagRaph)
    * d3Network: [networkD3](http://christophergandrud.github.io/networkD3/)
    * Circle plot: [edgebundleR](https://github.com/garthtarr/edgebundleR)
    * sigma: [sigma](https://github.com/jjallaire/sigma), 
    * hive: [d3hiveR](http://www.buildingwidgets.com/blog/2015/7/11/week-27-d3hiver)
* `ggvis`
* `googleVis`
* Server based interactivity - `shiny`
